---
title: "Montgomery County Crash Data"
author: "Alexander Morales & Andrea Cano"
date: "March 13th, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This document visualizes crash data in the Montgomery County. It provides graphical evidence to answer multiple questions about safety or just simple curiosities when it comes to crashes.

```{r}
dat=read.csv(file = "Crash_report.csv")
```

## Where Data got collected and why ?

This document explores the information provided on the crash reports in Montgomery County. This information is focused on the traffic collisions on local roadways collected by the Automated Crash Reporting System (ACRS) from the Maryland State Police.

## Data exploration

```{r}
#imputing N/A data to UNKNOWN
#plotting weather
#dat$Weather[dat$Weather == "N/A"] = NA
weather = dat$Weather
weather = gsub("N/A","UNKNOWN", weather) 
barplot(sort(table(weather)),horiz = TRUE, las=1,col = "lightblue", main = "Weather During Crash")

#checking Seasons
dat$dtime = strptime(as.character(dat$Crash.Date.Time),"%m/%d/%Y %I:%M:%S %p", tz ="EST")
months = dat$dtime$mon +1
months = month.abb[months]
barplot(sort(table(months)), horiz = TRUE, las = 1, main = "Crashes per Month", col = "blue")

#Plotting time of crashes 

library(ggmap)
#separating lat and long from location
coordinate = gsub("[()]","",dat$Location)
coordinate = unique(sort(coordinate))
map = get_map(coordinate[1], maptype = "satellite",zoom = 16)
p = ggmap(map)
p





```
## Data cleaning and preprocessing
```{r}

substance=dat$Driver.Substance.Abuse
substance = gsub("N/A","UNKNOWN", substance) 
par(mar=c(3,12,3,3))
barplot(sort(table(substance)),horiz = TRUE,las=1,col = "lightblue", main = "Substance use",cex.names = .75,cex.axis = .75)

```
The data collected from Montegomery county had some "N/A" values for substance use.However; we were curious to see how many people were under the influence while conducting the vehicle in the county. We were astonished once we imputed all the "N/A" values in the data to "UNKNOWN" that majority of the accidents the driver was not under the influence. Also, that there was vast amount of "UNKNOWN", that could lead us to assume that the Police Department was not checking a large quantity of drivers if they were impaired with any alcohol or drugs. 
## Data exploration and visualization
```{r}


par(mar=c(3,12,3,3))
barplot(sort(table(dat$Vehicle.Movement)),horiz = TRUE,las=1,col = "lightblue", main = "Vehicle Movement",cex.names = .75,cex.axis = .75)


t.control=dat$Traffic.Control
t.control = gsub("N/A","UNKNOWN",t.control) 
par(mar=c(3,12,3,3))
barplot(sort(table(t.control)),horiz = TRUE,las=1,col = "lightblue", main = "Traffic Control",cex.names = .75,cex.axis = .75)

#plot(table(dat$Vehicle.Movement),type="p")
#agg= aggregate(Vehicle.Movement~Traffic.Control,data=dat,median)
dat$dtime = strptime(as.character(dat$Crash.Date.Time),"%m/%d/%Y %I:%M:%S %p", tz ="EST")
par(mar=c(3,3,.5,.5))
hours = dat$dtime$hour + 1
tab = as.data.frame(table(hours))
#zscale this shoop
plot(table(hours), las = 1, main = "Hours", col = "blue",type = "l")
dtab = as.data.frame(table(hours))
maH = names(which.max(table(hours)))
max.line = as.numeric(maH)
abline(v = max.line, col = "red")
miH = names(which.min(table(hours)))
min.line = as.numeric(miH)
abline(v = min.line, col = "green")



```
The graph above shows the analysis of the movement of the vehicles that was a factor for their car accident. We can conclude that a vast amount of accidents occured while the vehicle was moving at a constant speed. However; if the most of the accidents are caused by constant speed then we could assume that the satelite image of the road with the most accidents is caused by speed.  


The barplot above shows the traffic control in the regions where all the accidents occured.It is highly noticeable that most accidents occur when their isn't traffic control on the roads, according to the barplot. Also, most of the accidents occur on the road shown on the satellite image above. Therefore, we can safely assume that the road needs traffic control in order to lower the rate of accidents on that road. 



## Conclusion
To conclude our data exploration of the crash data of Montgomery county in Maryland. After cleaning several factors that came into mind for the high volume of accidents in the county. We could safely assume that the majority of the accidents are caused by speeding, and the drivers in that region are not under the influence in most cases. We took into consideration the traffic control, and we concluded with a graph that most accidents occured when traffic controls are not present. Therefore, in order to lower the accident rate in some of the roads in Montgomery county some traffic control will need to be inplaced in order to keep drivers from speeding through lonely roads. 